---
output: 
  rmarkdown::html_vignette:
    keep_md: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, message = FALSE}
library("openxlsx")
library("tidyverse")
```

First, read in the acoustics master spreadsheet downloaded from the
[project website](http://www.phonetics.ucla.edu/voiceproject/voice.html).

```{r}
acoustics <- read.xlsx("website/Spreadsheet/Voice_Master.xlsx")

glimpse(acoustics)
```

In row 6460 (6461 with header), the column names are repeated in place of the
measurements for the `shrF0` columns, so these columns were all read as type
`character`, and the values in this row are invalid. Coerce these columns to
type `double`, which will change the extra column names in that row to `NA`.

```{r}
acoustics <- mutate_at(acoustics, vars(starts_with("shrF0")), as.double)
```

The tone number, aspiration, and tone contour columns were read as type
`double`. Coerce these columns to the correct types.

```{r}
acoustics <- mutate(acoustics,
  No.Tone = as.integer(No.Tone),
  Aspiration = as.logical(Aspiration),
  Tone.Cont = as.character(Tone.Cont)
)
```

Discard the overall mean columns, which won't be used when the dataframe is in
the tidyverse format.

```{r}
acoustics <- select(acoustics, -ends_with("mean"))
```

The column named `0=do.not.use` has a `0` to indicate bad data, but sometimes
another annotation too:

```{r}
count(acoustics, `0=do.not.use`)
```

If this column indicates bad data for some or all of the acoustic measurements,
change those acoustic measurements to `NA`.

```{r}
beg_acoustics_cols <- str_subset(names(acoustics), "means(00[1-3])$")
mid_acoustics_cols <- str_subset(names(acoustics), "means(00[4-6])$")
end_acoustics_cols <- str_subset(names(acoustics), "means(00[7-9])$")
```

```{r}
acoustics[
  !is.na(acoustics$`0=do.not.use`) & acoustics$`0=do.not.use` == "0",
  c(beg_acoustics_cols, mid_acoustics_cols, end_acoustics_cols)
] <- NA

acoustics[
  !is.na(acoustics$`0=do.not.use`) & acoustics$`0=do.not.use` == "0 beg",
  beg_acoustics_cols
] <- NA

acoustics[
  !is.na(acoustics$`0=do.not.use`) & acoustics$`0=do.not.use` == "0 mid",
  mid_acoustics_cols
] <- NA

acoustics[
  !is.na(acoustics$`0=do.not.use`) &
    acoustics$`0=do.not.use` %in% c("0 end", "0 end ", "check end f0", "check f0 end"),
  end_acoustics_cols
] <- NA
```

There is also one duplicated filename, with different measurements for two
entries:

```{r}
filter(acoustics, Filename == "speaker3M_aC32a.mat")
```

Based on the original audio and the timestamps in the original TextGrid, the
second entry is incorrect, so remove it from the table.

Also, remove the `0=do.not.use` column, now that the bad data that was indicated
by this column has been converted to `NA` or removed.

```{r}
acoustics <- filter(acoustics,
  Filename != "speaker3M_aC32a.mat" | seg_End < 140
) %>%
  select(-`0=do.not.use`)
```

Remove the `.mat` extension from the rest of the filenames.

```{r}
acoustics <- mutate(acoustics, Filename = str_remove(Filename, "\\.mat$"))
```

Rename and recode some columns with the descriptive names from the PDF readme.

```{r}
acoustics <- mutate(acoustics,
  language_variety = recode(`Dialect/Village`,
    Black = "Black Miao",
    CH = "China",
    TW = "Taiwan",
    SJG = "San Juan Guelavia Zapotec",
    SMZ = "Santiago Matatlan Zapotec",
    v1 = "Village 1",
    v2 = "Village 2",
    W = "White Hmong"
  ),
  phonation = recode(Phonation,
    M = "Modal",
    C = "Creaky",
    B = "Breathy",
    L = "Lax",
    `T` = "Tense"
  ),
  vowel_nasality = recode(`Oral/Nasal`,
    O = "Oral",
    N = "Nasal"
  ),
  tone_category = recode(Tone,
    H = "High",
    M = "Mid",
    L = "Low",
    B = "Big"
  ),
  segment_type = recode(CorV,
    C = "Consonant",
    V = "Vowel"
  ),
  consonant_aspiration = ifelse(Aspiration,
    "Aspirated",
    "Unaspirated"
  ),
  consonant_nasality = recode(Pre_C,
    O = "Oral",
    N = "Nasal"
  )
)

acoustics <- select(acoustics,
  -`Dialect/Village`,
  -Phonation,
  -`Oral/Nasal`,
  -Tone,
  -CorV,
  -Aspiration,
  -Pre_C
)
```

Remove columns that don't provide any unique information.

```{r}
acoustics <- select(acoustics,
  -Speaker, # composite of Sex + Speaker_Number columns
  -Lphon,   # composite of Language + Phonation columns
  -Tphon    # composite of Tone + Phonation columns
)
```

Rename the columns in the tidyverse style.

```{r}
acoustics <- rename(acoustics,
  textgrid_label = Label,
  speaker_number = `Speaker.#`,
  speaker_id = Lang_Spk,
  speaker_sex = Sex,
  tone_contour = Tone.Cont,
  number_of_tone_targets = No.Tone,
  language_specific_tone = Tone.From.txtgrid,
  vowel_quality = Vowel
)

# lowercase, but avoid lowercasing the measurement columns, eg H1c
acoustics <- rename_if(acoustics, funs(!is_double(.)), tolower) %>%
  rename(segment_end = seg_End, segment_start = seg_Start, duration = Duration)
```

Reshape the dataframe. Columns ending in 001-009 are timepoints.

```{r}
acoustics <- acoustics %>%
  gather(key, value, matches("00\\d$")) %>%
  extract(key, c("measure", "timepoint"), "(.+?)_means00(\\d)") %>%
  spread(measure, value) %>%
  mutate(timepoint = as.integer(timepoint))
```

Rearrange the columns:

1. Information about the speaker and their language
2. Name of the audio file and the timestamps in the file where this segment
comes from
3. Information about the segment
4. Information about the preceding consonant, if applicable
5. Acoustic measures: pitch, formant filter, harmonic amplitudes, measures
derived from the harmonic amplitudes, energy and noise measures

```{r}
acoustics <- select(acoustics,
  speaker_id, speaker_number, speaker_sex, language, language_variety,
  filename, segment_start, segment_end, duration, timepoint,
  textgrid_label, segment_type, vowel_quality, phonation, vowel_nasality,
  tone_contour, number_of_tone_targets, tone_category, language_specific_tone,
  consonant_aspiration, consonant_nasality,
  strF0, sF0, pF0, shrF0,
  sF1, sB1, sF2, sB2, sF3, sB3, sF4, sB4, pF1, pF2, pF3, pF4,
  H1u, H1c, H2u, H2c, H4u, H4c, A1u, A1c, A2u, A2c, A3u, A3c,
  H1H2u, H1H2c, H2H4u, H2H4c, H1A1u, H1A1c, H1A2u, H1A2c, H1A3u, H1A3c,
  Energy, HNR05, HNR15, HNR25, HNR35, CPP, SHR
) %>%
  arrange(filename, segment_start, timepoint)
```

In the reshaped table, individual measures with bad data at a particular
timepoint are marked with a `0` placeholder instead of a measurement. Change
these to `NA`.

However, some measures can also have valid measurements that are exactly `0`.
For example, `H1A1u` is the difference in amplitude between the first harmonic
and the harmonic that is closest to the first formant. If these happen to be the
same harmonic, then `H1A1u` (and `H1A1c`) will be exactly `0`, and similarly for
the measures derived from `A2` and `A3`.

To fix these measures, re-calculate all of the measures derived from `A1`, `A2`,
and `A3` only for the cells that are now `NA`. All other `NA` values will
propagate correctly.

Also, `SHR` may have a true value of exactly `0` if there is zero estimated
subharmonic energy. Because this is a very common situation, there is no way to
determine if a `0` in this column is a true `0` measurement or if it indicates
bad data. Instead of replacing `0` values in this column with `NA`, make a note
in the documentation.

```{r}
acoustics <- mutate_at(acoustics, vars(strF0:CPP), funs(na_if(., 0))) %>%
  mutate(
    H1A1u = ifelse(is.na(H1A1u), H1u - A1u, H1A1u),
    H1A1c = ifelse(is.na(H1A1c), H1c - A1c, H1A1c),
    H1A2u = ifelse(is.na(H1A2u), H1u - A2u, H1A2u),
    H1A2c = ifelse(is.na(H1A2c), H1c - A2c, H1A2c),
    H1A3u = ifelse(is.na(H1A3u), H1u - A3u, H1A3u),
    H1A3c = ifelse(is.na(H1A3c), H1c - A3c, H1A3c)
  )
```

Save the data.

```{r, results = 'hide'}
usethis::use_data(acoustics)

write_csv(acoustics, path = "csv/acoustics.csv")
```
