---
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{gujarati}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette uses the [`voicequality`
package](https://github.com/scjs/ucla-voice-quality) to replicate two of the
figures in this paper:

> Khan, S. (2012). The phonetics of contrastive phonation in Gujarati. _Journal
of Phonetics_, 40(6), 780–795. https://doi.org/10.1016/j.wocn.2012.07.001

# Setup

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

First, load the `tidyverse` packages and the data from the `voicequality`
package.

```{r, message = FALSE}
library("tidyverse")

data("acoustics", package = "voicequality")
data("egg", package = "voicequality")
```

For the Gujarati part of the [UCLA Voice Quality
Project](http://www.phonetics.ucla.edu/voiceproject/voice.html), Sameer Khan
recorded ten speakers producing a list of words. There is one Gujarati-language
audio file per speaker for each target word. Each file has several repetitions
of a sentence that includes the word. One target vowel was extracted from each
repetition, and acoustic and EGG measurements for all of the vowel tokens across
all of the speakers and words were collected in the `acoustics` and `egg`
tables.

Subset the Gujarati data from the `acoustics` table and join it with the `egg`
table.

```{r, message = FALSE}
gujarati <- filter(acoustics, language == "Gujarati") %>%
  left_join(egg)
```

List all of the columns in the data. To read the documentation for these
columns, use `help("acoustics", package = "voicequality")` and `help("egg",
package = "voicequality")`.

```{r}
names(gujarati)
```

Show a few of the columns.

```{r}
select(gujarati, speaker_id, filename, segment_start, timepoint, CQ_H, H1H2c) %>%
  as_tibble()
```

Each vowel token can be uniquely identified by its combination of `filename` and
`segment_start` values. Each vowel token was divided into nine equal windows of
time. This table has a row of measurements for each timepoint, so it includes
nine rows per vowel token. Each row of measurements can be uniquely identified
by its combination of `filename`, `segment_start`, and `timepoint` values.

Each Gujarati word has a file number assigned to it. The last two digits in each
filename give the file number for the word that was recorded in that file.

Extract the file numbers from each filename into a separate column.

```{r}
gujarati <- extract(gujarati,
  filename, "file_number", "^.*(\\d{2})_Audio$", remove = FALSE
)
```

Khan (2012) used a subset of the Gujarati data in the `voicequality` package.
The appendix has a list of the words that were used for the paper. The file
numbers for those words in the `gujarati_word_list` table are:

```{r}
khan_2012_stimuli <- tibble(
  file_number = c(
    "01", "03", "06", "08", "09", "11", "12", "13", "30", "31", "32", "33",
    "34", "35", "38", "39", "40", "41", "46", "47", "48", "49", "56", "57",
    "59", "60"
  ),
  vowel_group = c(
    "a", "a", "a", "a", "a", "a", "a", "a", "e", "e", "e", "e",
    "e", "e", "e", "e", "e", "e", "e", "e", "e", "e", "o", "o",
    "o", "o"
  )
)
```

Join the complete Gujarati data with the list of file numbers to get the subset
used in Khan (2012).

```{r, message = FALSE}
khan_2012 <- inner_join(gujarati, khan_2012_stimuli)
```

Reorder the levels of the `phonation` variable so that they will be in the
correct order for the graph.

```{r}
khan_2012 <- mutate(khan_2012,
    phonation = fct_relevel(phonation, c("Modal", "Breathy"))
)
```

# Timecourse of contact quotient measure (Figure 10)

The left panel of Figure 10 in Khan (2012) compares the timecourses of the
contact quotient for modal versus breathy vowels.

The `voicequality` data has measurements at nine timepoints within each vowel
token. This figure plots the mean and standard error of per-speaker averages for
the two phonation types at each timepoint. To recreate the figure, first
calculate the average contact quotient for each phonation type, for each
speaker. Speakers are uniquely identified by the `speaker_id` column.

```{r}
mean_cq <- khan_2012 %>%
  group_by(speaker_id, timepoint, phonation) %>%
  summarize(CQ_H = mean(CQ_H, na.rm = TRUE)) %>%
  ungroup()

mean_cq
```

Next, plot the means and standard errors. `ggplot2` can calculate these itself
before plotting by using the `summary` stat with the appropriate functions.

```{r}
ggplot(mean_cq, aes(timepoint, CQ_H, shape = phonation, linetype = phonation)) +
  geom_point(stat = "summary", fun.y = mean) +
  geom_line(stat = "summary", fun.y = mean) +
  geom_errorbar(stat = "summary", fun.data = mean_se, width = 0.1) +
  scale_linetype_manual(values = c("Modal" = "solid", "Breathy" = "longdash")) +
  scale_x_continuous(breaks = seq(1, 9)) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

The results look very close to the left panel of the published Figure 10, except
that all of the values are shifted up by about 0.05.

# Average spectral tilt measures (Figure 4)

Figure 4 in Khan (2012) shows the average values of four spectral tilt measures
for modal versus breathy vowels in each of the three vowel groups.

To recreate this figure, first calculate the average spectral tilt for each
phonetic category for each speaker. There are six phonetic categories, with one
category for each combination of `phonation` (`Modal` or `Breathy`) and
`vowel_group` (`a`, `e`, or `o`). Speakers are uniquely identified by the
`speaker_id` column. Note that this calculation averages over all nine
timepoints in each vowel token.

```{r}
mean_spectral_tilt <- khan_2012 %>%
  group_by(speaker_id, vowel_group, phonation) %>%
  summarize(
    H1H2c = mean(H1H2c, na.rm = TRUE),
    H2H4c = mean(H2H4c, na.rm = TRUE),
    H1A1c = mean(H1A1c, na.rm = TRUE),
    H1A3c = mean(H1A3c, na.rm = TRUE)
  )

mean_spectral_tilt
```

One way to make a single four-panel plot is to `gather()` the variables to be
plotted, then plot each one in a separate facet.

```{r}
mean_spectral_tilt <- mean_spectral_tilt %>%
  gather(measure, measurement, H1H2c, H2H4c, H1A1c, H1A3c) %>%
  mutate(measure = fct_relevel(measure, c("H1H2c", "H2H4c", "H1A1c", "H1A3c")))
```

```{r, fig.height = 6, fig.width = 6}
ggplot(
  data = mean_spectral_tilt,
  mapping = aes(
    x = vowel_group,
    y = measurement,
    fill = phonation
    )
  ) +
  geom_bar(
    stat = "summary",
    fun.y = mean,
    position = position_dodge(0.9)
  ) +
  geom_errorbar(
    stat = "summary",
    fun.data = mean_se,
    position = position_dodge(0.9),
    width = 0.1
  ) +
  facet_wrap(~measure, scales = "free") +
  theme_minimal() +
  theme(
    axis.title = element_blank(),
    legend.position = "bottom"
  )
```

The resulting values look similar to the published Figure 4.
